\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-06 tex/paper.tex
some non-final notes on other works

As thoroughly explored in \citep{2017arXiv170104862A}, GANs remain remarkably
difficult to train because of vanishing gradients as $d$ saturates, or because of
unreliable updates when the training procedure is relaxed. As a remedy,
Wasserstein GANs~\citep{2017arXiv170107875A} reformulate the adversarial
setup in order to minimize the Wasserstein-1 distance $W(p_r(\mathbf{x}), p(\mathbf{x}|\theta))$ by
replacing the adversarial classifier with a 1-Lipschitz adversarial critic
$d(\cdot; \phi) : \mathbb{R}^d \to \mathbb{R}$. 
\begin{align}
   W( p_r(\mathbf{x} ), p(\mathbf{x} | \theta)  ) = \mathcal{L}_W = \sup_{f} \mathbb{E}_{p(\mathbf{x}|\theta)} [f(\tilde{\mathbf{x}})] - \mathbb{E}_{p_r(\mathbf{x})} [f(\mathbf{x})]
\end{align}
Under the WGAN-GP formulation of \cite{2017arXiv170400028G}
for stabilizing the optimization procedure,
training $d$ and $g$ results in alternating gradient updates on $\phi$ and $\theta$ in order to respectively minimize

@@

Variational optimization~\cite{2012arXiv1212.4507S} and evolution strategies~\citep{2011arXiv1106.4487W} are general
optimization techniques that can be used to form a differentiable bound
on the optima of a non-differentiable function. Given a function $f$ to minimize,
these techniques are based on the simple fact that
\begin{equation}
    \min_{\theta \in \Theta} f(\theta) \leq \mathbb{E}_{\theta \sim q(\theta|\psi)} [f(\theta)] = U(\psi),
\end{equation}
where $q(\theta|\psi)$ is a proposal distribution with parameters $\psi$ over input values $\theta$.
That is, the minimum of a set of function values is always less than or equal
to any of their average. Provided that the proposal is flexible enough, the parameters $\psi$
can be updated to place its mass arbitrarily tight around the optimum $\theta^* = \min_{\theta \in \Theta} f(\theta)$.

@@

Original OPVI
\begin{eqnarray}
\lambda^* = \inf_\lambda \sup_\phi \mathbb{E}_{ z \sim q( z | \lambda)} [ (O^{p,q} f_\lambda) ] 
\end{eqnarray}
with
\begin{equation}
(O^{p,q} f) = \log q(z) - \log p(z|x) \, \forall f \in \mathcal{F}
\end{equation}

@@

\glnote{Discuss 1702.08235 and references therein on adversarial variational bayes/inference.}

@@

>>>>

As thoroughly explored in \citep{2017arXiv170104862A}, GANs remain remarkably
difficult to train because of vanishing gradients as $d$ saturates, or because of
unreliable updates when the training procedure is relaxed. As a remedy,
Wasserstein GANs~\citep{2017arXiv170107875A} reformulate the adversarial
setup in order to minimize the Wasserstein-1 distance $W(p_r(\mathbf{x}), p(\mathbf{x}|\theta))$ by
replacing the adversarial classifier with a 1-Lipschitz adversarial critic
$d(\cdot; \phi) : \mathbb{R}^d \to \mathbb{R}$. 
\begin{align}
   W( p_r(\mathbf{x} ), p(\mathbf{x} | \theta)  ) = \mathcal{L}_W = \sup_{f} \mathbb{E}_{p(\tilde{\mathbf{x}}|\theta)} [f(\tilde{\mathbf{x}})] - \mathbb{E}_{p_r(\mathbf{x})} [f(\mathbf{x})]
\end{align}
Under the WGAN-GP formulation of \cite{2017arXiv170400028G}
for stabilizing the optimization procedure,
training $d$ and $g$ results in alternating gradient updates on $\phi$ and $\theta$ in order to respectively minimize

@@

Variational optimization~\cite{2012arXiv1212.4507S,staines2013optimization} and evolution strategies~\citep{2011arXiv1106.4487W} are general
optimization techniques that can be used to form a differentiable bound
on the optima of a non-differentiable function. Given a function $f$ to minimize,
these techniques are based on the simple fact that
\begin{equation}
    \min_{\theta \in \Theta} f(\theta) \leq \mathbb{E}_{\theta \sim q(\theta|\psi)} [f(\theta)] = U(\psi),
\end{equation}
where $q(\theta|\psi)$ is a proposal distribution with parameters $\psi$ over input values $\theta$.
That is, the minimum of a set of function values is always less than or equal
to any of their average. Provided that the proposal is flexible enough, the parameters $\psi$
can be updated to place its mass arbitrarily tight around the optimum $\theta^* = \min_{\theta \in \Theta} f(\theta)$.

@@

Original OPVI  \citep{2016arXiv161009033R}

\begin{eqnarray}
\lambda^* = \inf_\lambda \sup_\phi \mathbb{E}_{ z \sim q( z | \lambda)} [ (O^{p,q} f_\lambda) ] 
\end{eqnarray}
with
\begin{equation}
(O^{p,q} f) = \log q(z) - \log p(z|x) \, \forall f \in \mathcal{F}
\end{equation}

@@

This work has many connections to variational inference, in which the goal is to optimize the recognition model $q(\theta|\psi)$ so that it is close to the true posterior $p(\theta|x)$. This is slighty different from variational optimization (or variational learning), where the toal is to find $\theta^*$ such that $p(x|\theta^*) \approx p_r(x)$.  Typically in variational inference the model $p(x|\theta)$ admits a density, so that the likelihood is known. Thus, assuming the prior also admits a density, it is tractable to evaluate the posterior is (up to a constant factor). In Adversarial Variational Bayes \citep{DBLP:journals/corr/MeschederNG17} and , equivalently, the \texttt{PC-Adv} algorithm of  \citep{2017arXiv170208235H}), the authors use an adversarial setup, but assume a known density $p(x|\theta)$ that is differentiable with respect to $\theta$.

In \citep{tran2017variational}, the authors consider Variational Bayes with an intractable likelihood. In that approach "The only requirement is that the intractable likelihood can be estimated unbiasedly.'' In the case of most simulators it is not clear how one would obtain the unbiased estimator $\hat{p}(x|\theta)$.

In \citep{2017arXiv170208235H} the author defines implicit distributions to be ``probability models whose probability density function may be intractable, but there is a way to 1) sample from them exactly and/or calculate and approximate expectations under them, and 2) calculate or estimate gradients of such expectations with respect to model parameters.'' This notion of implicit model does not include most scientific simulators as 2) is not satisfied. The  \texttt{JC-Adv} and \texttt{JC-Den} algorithm defined there applies to implicit models, but, as the author points out, is aimed at variational inference and ``does not provide a direct way to learn model parameters $\theta$''.

@@

\textbf{KC: could also use this to tune non-differentialble fast simulator to slower full simulator. Then fixed data size not an issue}
